---
title: "Value of Forecasting in Emergency Service Operational Planning"
subtitle: ""
author: "Bahman Rostami Tabar"
#date: '25/11/2020'xaringan::inf_mr()
output:
  xaringan::moon_reader:
    chakra: ./libs/remark-latest.min.js
    css: ["default","././libs/slides.css"]
    lib_dir: ./libs
    seal: false
    anchor_sections: false
    nature:
      highlightStyle: github
      highlightLanguage: ["r","markdown","yaml"]
      ratio: '16:9'
      highlightLines: true
      countIncrementalSlides: false
      navigation:
        scroll: false 
---
class: inverse
background-image:  linear-gradient(to right, rgba(50, 50, 50, .5), rgba(50, 50, 50, .4)),url("resources/titlepage.png")
background-size: cover

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, width = 120)

library(tidyverse)
library(knitr)
library(kableExtra)
library(fontawesome)
library(flair)
library(DT)
library(gapminder)
library(ggthemes)
library(countdown)

opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
  fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center',
  out.width = "70%", fig.show = 'hold',
  cache = TRUE, external = TRUE, dev = 'svglite', dev.args = list(bg = "transparent")
)

#fig.width = 12, fig.height = 4

mp4_vid <- function(src){
  HTML(
    paste0(
      '<video autoplay>
        <source src="', src, '" type="video/mp4">
      </video>'
    )
  )
}


theme_set(
  theme_grey(base_size = 16) +
  theme(
    legend.position = "bottom",
    plot.background = element_rect(fill = "transparent"),
    legend.background = element_rect(fill = "transparent")
  )
)
```

```{r load-data, include=FALSE}

```

```{css, echo=FALSE}
/* custom.css */
.left-code {
  color: #777;
  width: 42%;
  height: 92%;
  float: left;
}
.right-plot {
  width: 56%;
  float: right;
  padding-left: 1%;
}
```


.title[Value of Forecasting in Emergency Service Operational Planning]
<!-- .sticker-float[![fable](resources/fable.svg)] -->



.bottom[
## Bahman Rostami-Tabar (`r fa("twitter", fill="#1da1f2")`[@Bahman_R_T](https://twitter.com/Bahman_R_T)) 
### Website [www.bahmanrt.com](https://www.bahmanrt.com/)
<!-- ### Slides @ [slides.mitchelloharawild.com/nhs2020](https://slides.mitchelloharawild.com/nhs2020) -->
### Quarterly Forecasting Forum
### 21 May 2021
]

---

## Outline
<hr>

- How forecasts can inform planning in Accident & Emergency department?

- Data & forecasting models

- Evaluate forecast accuracy using statistical measures

- Scheduling algorithm and utility measures

- What would be the performance of planning if we plan based on forecasts?

- What if we don't  bother with forecasting methods?

---
class: inverse, middle
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .6)), url("resources/as-staff.jpg")
background-size: cover

.larger[How many nurses/doctors and When?]

--

- .mlarger[Demand/capacity planning]
- .mlarger[Staffing]
- .mlarger[Rostering]


---
## Forcasts can inform various planning horizons

- 6 months

- 42 days

- 24/48 hours


---
## Forcasts can inform various planning horizons

While the strategic planning problems of EDs are more concerned with decisions such as how many rooms an ED should provide, the tactical decisions need to absorb demand uncertainty by providing effective and efficient staffing decisions.

---
##

- In this study, we focus on developing a planning approach: 

      - support the hospital management in strategic and tactical cost-efficient resource capacity dimensioning decisions.

- is able to provide a cost-sensitive evaluation of different forecasting methods.

---
## Model

We model the problem as a mixed-integer program which creates workforce schedules for internists, nurses and surgeons.
		%\item Simultaneously, hospital equipment such as reanimation rooms, examination and treatment rooms are assumed to be continuously available.
		%\item The model includes flexible working times for staff and therefore determines the optimal quantity of personnel and material resources while minimizing costs.
		
ED resources such as nurses as well as material resources such as rooms have to cover the demand. Since demand is fluctuating over the week, overtime costs can occur for human resources which, however, can be compensated within the week, or between several weeks.

---
class: inverse, top
background-image: linear-gradient(to right, rgba(300, 300, 300, .4), rgba(300, 300, 300, .5)), url("resources/questions.jpg")
background-size: cover

.hand-large[Thank you]

.remember[.larger[Question?]]

`r fa("twitter", fill = "white")` Say hello: [@Bahman_R_T](https://twitter.com/Bahman_R_T)



